### 2022-07-13 manuscript.qmd
round off phyo confounding section

---
title: "Causal inference and co-evolutionary contingencies using dynamic phylogenetic models"
format: html
editor: visual
bibliography: references.bib
---

## Introduction

Phylogenetic comparative methods (PCMs) are commonly used to study the co-evolution of organismal traits, spanning topics such as anatomy and physiology [@garland2005phylogenetic; @oconnor2022immunity; @thayer2018impact; @navalon2019evolutionary; @dunn2015evolutionary], life history and behavior [@maclean2012does; @clayton1994comparative; @salguero2016fast; @bielby2007fast; @cornwallis2017cooperation; @wilkinson2019recurrent], and cultural evolution [@mace2005phylogenetic; @watts2015broad; @navarrete2016coevolution]. "Co-evolution" refers to repeatable patterns of trait co-variation over evolutionary time, across species or across populations. In practice, co-evolution can be investigated using a diverse family of statistical techniques, depending on the research question and type of data available [@harvey1991comparative; @garamszegi2014modern; @nunn2011comparative].

Trait co-evolution simply describes *how* traits change over time. Such descriptions have value for understanding natural history, ancestral state reconstruction (using the present to retrodict the past), and can stimulate new research questions. However, simply inferring that traits changed together in the past does not tell us *why* they changed, or how the traits would respond to intervention. In short, evolutionary correlation does not imply evolutionary causation. Nonetheless, researchers in ecology and evolution are often motivated to study trait co-evolution because they believe these patterns reflect causal relations [CITATION/EXAMPLE HERE?]. While some authors use paradigms such as phylogenetic path analysis [POINT TO RELEVANT SECTION BELOW?] to justify their models and make their causal assumptions explicit, many others refrain from stating their assumptions at all, despite drawing implicit causal conclusions. Ambiguity about causal inference can lead to misleading claims and contradictory results, undermining the perceived value and reliability of comparative approaches in biology. [I WOULD MAKE THIS PARAGRAPH SLIGHTLY LESS 'OFFENSIVE', BY 1) STATING THAT THESE ISSUES ARE BY NO MEANS LIMITED TO PCM, AND 2) GIVE SOME EXAMPLES OF PCMS  THAT DO OR DO NOT DRAW CAUSAL CONCLUSIONS. OTHERWISE WE RISK UPSETTING READERS WHO EITHER DON'T THINK THIS IS A PROBLEM, OR DON'T LIKE TO BE SINGLED OUT AS PROBLEMATIC - e.g. ISLER & VAN SCHAIK EVO ANTHRO IS A GOOD EXAMPLE OF BEING COGNESCENT OF THE DIFFICULTY OF CAUSAL INFERENCE IN A PGLS/IC WORLD]

A major barrier to incorporating causal inference into study design, analysis, and interpretation is that the causal assumptions baked into popular PCMs are not always apparent to the end users (researchers). For example, the most widely used model of trait evolution in PCMs (Brownian motion) assumes a purely neutral model of evolution with a constant rate of drift over the entire phylogenetic tree. Another example is that, for virtually all models of the co-evolution of continuous traits, effects can only be interpreted causally if we assume that we have the direction of causality right (X -> Y, not Y -> X), and that there is no reciprocal causation.

In this article, we first review some popular classes of PCM and their assumptions and limitations with respect to causal inference. These range from standard regression models that lack clear causal interpretations over phylogenetic path analysis, which makes causal assumptions explicit, to dynamic models that detect temporal contingency (but are limited to binary traits). We then introduce a new dynamic method which overcomes some of these limitations and enables the study of directional trait co-evolution for an arbitrary number of traits that can be used with any type of data (continuous, discrete, ordered).

## Co-evolution as correlated drift: the evolutionary regression coefficient

The most common way to quantify co-evolution between traits is the evolutionary regression coefficient [@pagel1993seeking]. The evolutionary regression coefficient is a measure of association between traits over evolutionary time, $b = \frac{\sigma(X,Y)}{\sigma^2(X)}$. It can be interpreted as "a one unit increase in trait $X$ is associated with a $b$ unit change in trait $Y$." When can we interpret these statistics as causal effects? The short answer is...almost never! To explain why, we sketch out assumptions needed to map an evolutionary regression coefficient onto a causal estimand.

First, what do we mean by causal? Following Pearl (some citations), we say that $X$ is a cause of $Y$ when $P(Y|do(X = x)) \neq P(Y|do(X = \textrm{not x}))$. The do($\cdot$) operator represents an *intervention* in which the value of some variable is fixed (for example, through assignment in a randomized experiment). This is an inclusive definition, where if the probability distribution $P(Y)$ is sensitive to the value of $X$ that is assigned, then $X$ has a causal effect on $Y$. Note here that $X$ refers to a random variable while $x$ refers to some realized unit value of that variable. In any observational study, we cannot directly access $P(Y|do(X = x))$. Instead, we have just have the condition probability of $Y$ given $X$, $P(Y|X)$. Pearl and colleagues' "do calculus" is a set of rules that tells us under what conditions $P(Y|do(X)) = P(Y|X)$, or when we can make them equivalent by adjustment for other variables $P(Y|do(X),Z) = P(Y|X,Z)$. An *estimand* is a unit specific estimate such as "what would we expect species $i$ trait value $y_i$ to be if $x_i$ took on a given value?". Under a particular set of causal and parameteric assumptions, a regression coefficient $b$ can be interpeted as the estimand "what is the difference in the expected (mean) value of $Y$ if $x = 1$ vs $x = 0~$?", or $\mathbb{E}[Y|x=1] - \mathbb{E}[Y|x=0]$.

An unitiuitive aspect of phylogenetic comparative methods is that cross-sections of species traits are used as an estimate of the *change* in $Y$ in response $X$. After all, we have only observed current trait values, not their evolution over time. It is unituitive not because it is incorrect in principle, but because it requires the reseracher to make a series of strong assumptions that may be in conflict with the stated research question. Following [@pagel1993seeking], we define a causal model of trait co-evolution in which we would be justified in interpreting the evolutionary regression coefficient as a causal effect. The model is defined below and depicted graphically in Figure 1.

Given a time series of trait evolution with $T$ time points, $\forall ~t \in T$

$$ x_t = x_{t-1} + \epsilon_{x,t}$$
$$ y_t = y_{t-1} + b(x_t - x_{t-1}) + \epsilon_{y,t} = y_{t-1} + b\epsilon_{x,t} + \epsilon_{y,t} $$
$$ \epsilon_{x,t} \sim \mathcal{N}(0,\sigma_x^2) $$
$$ \epsilon_{y,t} \sim \mathcal{N}(0,\sigma_y^2) $$
Expressed as discrete time difference equations:

$$ \Delta x_t =  \epsilon_{x,t}$$
$$ \Delta y_t = b\epsilon_{x,t} + \epsilon_{y,t}$$
This model assumes that all change in trait $X$ is due to random drift $\epsilon_x$. Likewise, change in trait $Y$ is due to independent drift $\epsilon_y$ and response to *changes* in $X$. There is no directional change over time, and the evolution of both traits is state-independent (i.e., $Y$ has an invariant response to drift in $X$, regardless of how high or low it is). We can relax some assumptions, such as constant $\sigma$ and $\b$ over time, and still get unbiased estimates of $b$ using a cross-sectional regression of species. Additionally, if $Y$ has a non-linear response to $\epsilon_x$, then $b$ will still be a valid estimate of the linear component--albeit now possibly misleading if the total response exhibits strong non-linearity. But if either trait is under selection (i.e., the edges between $Y_t$ and $Y_{t+1}$ or between $X_t$ and $X_{t+1}$ aren't = 1), then the regression coefficient will no longer be equal to b. To get a sense for why, imagine that, in addition to drift, $X$ is under moderate stabilizing selection. This means that $X_t$ is no longer just a record of the additive evolutionary history of $\epsilon_x$, and that a causal effect of $X$ on $Y$ is not evident from the trait covariance at the tips of a phylogeny. Just because you can't see it anymore, doesn't mean it isn't important.

![Figure 1](figures/MA_DAG.png){width=60%}

We previously stated that a cross section of current species trait values can provide an unbiased estimate of the change in trait $Y$ in response to changes in $X$. This fact is well known in the comparative literature [@pagel1993seeking; @rohle2006comment], and consistent with the rules of do-calculus--given the strict assumptions of the model outlined before. However, if traits were correlated in the past but no longer exhibit any association, then evolutionary history acts as a confounder. For example, in Figure 4B, $U_0$ creates a "backdoor path" that connects all pairs of $X_t$ and $Y_t$. This means that $P(Y|do(X)) \neq P(Y|X)$, unless we can somehow adjust for $U_0$.

The confounder denoted $U$ is unobserved, meaning that we can't statistically adjust for it. However, note that for any time $t$, $X_t$ and $Y_t$ are independent of $U_0$ conditional on $\{ X_{t-1},X_{t-1}$. In other words, the present is conditionally independent of $U_0$ (or "d-seperated") given the past. But in most comparative studies, we have not observed the past, only the present. This is where phylogeny can help us, because more closely related species or populations have more similar confounding factors (e.g., shared genetic architecture, environmental similarity), on average, than distantly related ones. Assuming that we have an appropriate model of evolution, then the patristic distance between species (i.e., time since divergence from a common ancestor) allows us to probabalistically reconstruct unobserved past values of $X$ and $Y$, closing the backdoor path $X \leftarrow U \rightarrow Y$ (cite Rethinking 2nd edition). It should be noted that such a scenario is sometimes called "phylogenetic confounding," but...[DOESN"T WORK IF U IS STILL PRESENT AT THE TIPS. HOW CAN WE TELL IF ITS PHYLOGENETIC CONFOUNDING VS A CAUSAL RELATIONSHIP? IN THE FORMER, THE COVARIANCE BETWEEN X AND Y WILL DECREASE OVER TIME. IN THE LATTER IT IS CONSTANT/STABLE].

Phylogentic position is just a type of clustering, and the presence of clusters (such as a clade) in comparative datasets makes analyses more susceptible to Simpson's paradox, a reversal of a trend at the group level (clade) vs the individual level (species). Statistical adjustment for phylogeny helps reduce both the type I (false-positive) and type II (false-negative) error rates (plus the Type M and Type S errors [ADD CITATION FOR WHAT THESE ARE]). It should be noted here that "adjusting for phylogeny" is not a magic-fix for unobserved confounding that is persistent throughout the phylogenetic tree. But it can help us adjust for confounding in the past.

Additionally, even if there was zero structural phylogenetic confounding (Figure 1A), it would still be a good idea in any finite sample to adjust for phylogeny because, as a source of variation in $Y$, phlyogeny is a "good control" in that it helps us isolate variation in $Y$ that is due to possible causal effects of $X$. Marginalizing over phylogenetic position can also give us improved estimates of the average effects when clades are unequally sampled.

Now that we have a sense for the assumed causal model underlying most phylogenetic comparative analyses, and motivation for why we statistically adjust for phylogeny, let's review some popular methods for estimating the evolutionary regression coefficient and introduce some variations on the causal model in Figure 1.

### Phylogenetic Generalized Least Squares (PGLS)

Phylogenetic generalized least squares (PGLS) is a variant of linear regression that accounts for the covariance between species or populations due to shared evolutionary history (likewise for its predecessor, independent contrasts) [@grafen1989phylogenetic; @symonds2014primer; @blomberg2012independent]. Given some outcome $Y$ and a co-evolutionary predictor $X$, an observation $i$ from species or population $j$ is modeled as:

$$y_{ij} = \beta_0 + \beta_xx_{ij} + \epsilon_{ij}$$ $$ \epsilon \sim  \mathcal{N}(0, \boldsymbol{\Sigma})$$ $$\boldsymbol{\Sigma} = \sigma^2\boldsymbol{V} $$ Where $\mathcal{N}$ denotes a normal or Gaussian probability density function and $\sigma^2$ is the variance of $Y$ conditional on $X$. $\beta_x$ is an estimator of the evolutionary regression coefficient that takes into account the phylogenetic non-independence between tip species. The covariance between related species is proportional to the off-diagonals of the phylogenetic variance-covariance matrix <!--# isn't this also often called A? doesn't matter i guess.. --> $\boldsymbol{V}$. To construct $\boldsymbol{V}$, we need (1) a matrix of patristic distances between species (i.e., time since divergence from last common ancestor) and (2) an assumed model of trait evolution.

The most commonly assumed model is Brownian motion (also called the Wiener process and thus denoted $W$), where $W$ follows a random walk proportional to time ($T$), such that, for any time $t$:

$$ W_{T= t} - W_{T = 0} \sim\mathcal{N}(0,\sigma t) $$ Therefore, the variance for a given species or population $j$ is proportional to the total branch length of the tree leading from root to tip, and the covariance between $j$ and any other species or population $k$ is proportional to their amount of shared history. For example, imagine a phylogeny with only $J$ = 3 tip species. The last common ancestor of all species is dated to T = 1,000,000 years ago, and the last common ancestor of 2 and 3 is dated to 0.5T = 500,000 years ago. The implied phylogenetic variance-covariance matrix ($\boldsymbol{V}$) under a Brownian motion model is thus:

$$ \boldsymbol{V} = \sigma^2 \begin{bmatrix}
T, & 0, & 0 \\
0, & T, & 0.5T \\
0, & 0.5T, & T
\end{bmatrix} $$

Thus, the total variance for species 3 is the sum of its evolutionary history, $\sigma^2[t(2,3)] + \sigma^2[t(3) - t(2,3)]$ and the covariance between species 2 and 3 is equal to the sum of their *shared* evolutionary history $\sigma^2[t(2,3)]$ (the tree time for their most recent common ancestor), which is always some fraction of the species total evolutionary time $T_j$. If the tree is ultrametric (all tips equidistant from the root), $T_j$ = $T ~\forall ~j \in J$.

$\sigma$ is easier to interpret when we first scale all times by the total tree length $T$. We will denote such a scaled variance-covariance matrix as $\boldsymbol{V}_s$:

$$ \boldsymbol{V}_s = \frac{\boldsymbol{V}}{T} = \sigma^2 \begin{bmatrix}
1, & 0, & 0 \\
0, & 1, & 0.5 \\
0, & 0.5, & 1
\end{bmatrix} $$

Empirically, the actual degree of covariance between related species or populations may be less than expected under a pure Brownian motion model. For example, when evolution is rapid (or more mundanely, when measurement error on trait data is substantial) the "phylogenetic signal" will be weak such that even closely related taxa may have dramatically different trait values [@blomberg2003testing; @kamilar2013phylogenetic]. PGLS can estimate the strength of phylogenetic signal from comparative data, multiplying the off-diagonals of the variance-covariance matrix by $\lambda \in [0,1]$ [@pagel1999inferring]. For example:

$$ \boldsymbol{V_s} = \sigma^2 \begin{bmatrix}
1, & 0, & 0 \\
0, & 1, & \lambda0.5 \\
0, & \lambda0.5, & 1
\end{bmatrix} $$

This scaling is often referred to as a "branch length transformation," because shrinking the off-diagonals (shared evolutionary history) while holding the diagonals constant (independent evolution) is equivalent to "stretching" the terminal branches leading to the tips and "shrinking" the past co-evolutionary branches. $\lambda = 1$ indicates that the phylogenetic strength is as strong as expected under a Brownian motion model, while lower values means that a lower proportion of the trait variance can be attributed to shared evolutionary history. At the extreme of $\lambda = 0$, $\boldsymbol{V}_s$ becomes an identity matrix and PGLS reduces to standard least squares regression.

Brownian motion is the simplest and most widely used model, and often interpreted as "drift" across the phylogenetic tree. Other popular choices include Ornstein-Uhlenbeck, where change is due to both random drift and "stabilizing selection" around some optimal trait value, as well as Early Burst models of adaptive radiation where the rate of change ($\sigma$) decreases over time. NOTE ON OU COVARIANCE: An OU model where the response is undergoing adaptive evolution and the predictor is drifting also works. But if the predictor is under selection, things get messed up. In general, $\boldsymbol{V}$ is constructed by inputting a matrix of phylogenetic distances to some covariance function $C(x,x')$ such that $V_{[j,k]} = C(j,k)$.

Although PGLS is widely used in comparative biology, it has several key limitations. These include the assumption of Gaussian errors, proneness to overfitting due to a lack of parameter regularization, and failure to accommodate many common data structures such as repeated measures of species trait values, missing data, and measurement error. In the next section we review more flexible regression approaches, as well as an important alternative method of estimating evolutionary associations.

### Phylogenetic Generalized Multilevel Models

Why would we care about the phylogenetic correlation? Because this is an indicator that the traits co-evolve across the phylogenetic tree rather than just at the tips.

Phylogenetic generalized multilevel (or mixed-effects) models (PGLMMs) extend the study of trait coevolution by leveraging analytic strategies developed in quantitative genetics to accommodate more complex, multivariate model designs with non-Gaussian outcomes [@lynch1991methods; @hadfield2010general; @ives2010phylogenetic; @ives2011generalized]. Rather than explicitly scaling the off-diagonals of $\boldsymbol{V}_s$ with some parameter $\lambda$, phylogenetic signal is quantified as the ratio of the variance of phylogenetic random effects $\nu$ (also called 'mixed' or 'varying' effects) to total trait variance. For example, an observation of a Gaussian trait $Y$ from a species or population $j$ and predictor $X$ is modeled as:

$$ y_{i,j} = \beta_0 + \beta_{x} x_{ij} + \nu_{j} + \epsilon_{ij} \\$$ $$ \nu \sim \mathcal{N}(0,\boldsymbol{\Sigma}_{\nu}) \\$$ $$\epsilon_{ij} \sim \mathcal{N}(0,\boldsymbol{\Sigma}_{\epsilon}) \\$$ $$\boldsymbol{\Sigma}_{\nu} = \sigma^2_{\nu}\boldsymbol{V}_s \\$$ $$\boldsymbol{\Sigma}_{\epsilon} = \sigma^2_{\epsilon}\boldsymbol{I} \\$$

Where $\boldsymbol{I}$ is an identity matrix and thus the residuals $\epsilon$ are assumed conditionally independent. The "phylogenetic signal" is now estimated as $\lambda_{\textrm{PGLMM}} = \frac{\sigma_{\nu}^2}{\sigma_{\nu}^2 + \sigma_{\epsilon}^2}$ (adjusted for $X$) and equivalent to $\lambda_{\textrm{PGLS}}$, but with the greater flexibility afforded by PGLMMs. $\lambda_{\textrm{PGLMM}}$ is also sometimes referred to as "phylogenetic heritability" [@hadfield2010general], since it is calculated in the same way as heritability based on pedigrees in quantitative genetics.

In this model, $\beta_{x(\textrm{PGLMM})}$ has the same interpretation as $\beta_{x(\textrm{PGLS})}$. If the response was non-Gaussian, $\beta_{x(\textrm{PGLS})}$ has the same interpretation *on the latent scale*. Conversion of the phylogenetic regression coefficient $\beta_{x(\textrm{PGLMM})}$ to the observation scale requires inverting the link function, and will be sensitive to other model parameters.

### Mutli-response PGLMMs: contrasting macro and microevolution

Residual and phylogenetic correlations can be different! Why? Plasticity and/or microevolutionary response to X may not be the same as the macroevolutionary process, a violation of *ergodicity*. or correlated measurement error.

PHYLOGENETIC CORRELATION DIRECTLY RELATED TO evolutionary regression coefficient $b$.

$$ \rho(X,Y) = \frac{\sigma(X,Y)}{\sigma(x)\sigma(Y)} $$
$$ b = \rho(X,Y)\frac{\sigma(Y)}{\sigma(X)} $$

### Misc. Notes

OU ADAPATIVE EVOLUTION AREN'T JUST A FIX FOR VIOLATION OF BROWNIAN MOTION--THEY IMPLY AN ENTIRELY DIFFERENT THEORETICAL ESTIMAND! Difference in Y given difference in X vs difference in the equilibrium trait value of Y given X. In the latter, the mean is not ergodic?

OU BUTLER AND KING ALGO--PIECEWISE CONSTANT SELECTION REGIME. I THINK THIS MEANS THAT Y RESPONDS TO WHATEVER TRIAT VALUE X WAS AT THE START OF THE SEGMENT--DOESN'T RESPOND TO CHANGES IN X WITHIN THE SEG.

STATE DEPENDENT ANALYSES MAKE US MORE SUSCEPTIBLE TO DARWIN'S SCENARIO -- UNREPLICATED BURSTS. REMOVAL OF PHYLOGENETIC CONFOUNDING RELIES ON THE COVARIANCE BETWEEN X AND Y DECREASING OVER TIME. IF THEY ARE MAINTAINED BY SELECTION WITHIN A CLADE, WE MAY FOOL OURSELVES. (!!!). SOLUTION = CLADE SPECIFIC RANDOM EFFECTS.

TEST YOUR METHOD AGAINST MODEL MISSPECIFICATION--WHAT IF DATA WERE GENERATED UNDER CORRELATED DRIFT? AFTER CLARIFYING HOW THESE ARE FUNDAMENTALLY DIFFERENT QUESTIONS/ESTIMANDS, IMPORTANT TO SHOW HOW THEY INFUENCE EACH OTHERS ESTIMATES IF WE IGNORE THEM. MY INTUITION = IGNORING SELECTION LEADS TO BIAS, IGNORING DRIFT DOES NOT.

![Figure 1](figures/phylo_DAG.png)
### Phylogenetic Path Analysis

While the terminology is not always consistent in the literature, note that "phylogenetic correlation" refers to a . A residual correlation refers to the remaining trait covariance *after* accounting for similarity due to shared evolutionary history. The residual correlations derived from PCMs are often interpreted as evidence for convergent evolution, adjusted for statistical non-independence and confounding due to phylogeny.

Independent contrasts, PGLS, PGLMMs, PPA

Discrete data can be modeled in the GLMMM framework if we conceptualize discrete phenotypes (binary, ordinal, categorical) as manifestations of a latent continuous trait, as in the quantitative genetic threshold model [@wright1934; @felsenstein2005using].

## Dynamic state-space models of co-evolution

Pagel's discrete, Hidden Markov Models, Butler & King OU

can be identified statistically as non-zero between-trait phylogenetic correlations----or any other measure of mutual information over evolutionary time.

THREE SCENARIOS TO SIMULATE:

\(1\) Confounding by third variable X_3 -\> X_2, X_3 -\> X_1

\(2\) Correlated drift (CoV\[X_1,X_2\]) != 0)

\(3\) Including correlated drift (shrinking cross-effects)